---
title: "RSimplerEDA-vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{RSimplerEDA-vignette}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
# editor_options: 
#   markdown: 
#     wrap: 79
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Exploratory Data analysis (EDA) is an important step in any data analysis. However, carrying out EDA with the ggplot2 package requires a lot of coding effort. Moreover, it assumes a basic knowledge of functions and grammar of graphics syntax that are appropriate for visualizing categorical and numerical variables. The RSimplerEDA package addresses this issue by providing functions that are tailored to produce categorical, numerical and correlation plots using a single line of code. Furthermore, the package provides customization capability for the plots based on specific user needs (theme, title, font, size and etc.). The users are able to spend more time on analyzing the data set and less time configuring ggplot plot settings.

The RSimplerEDA contains the following three functions:

1.  `corr_map`: Plot a correlation map with the given dataframe object and a character vector with numerical features. Users are allowed to set multiple arguments regarding the setting of the correlation plot including method to calculate the correlation, color schemes, and plot title.

2.  `numerical_eda:` This function takes in a data frame object, two numeric columns, and produces either a scatter or line plot to visualize the relationship between the two numerical features. Users can optionally change default arguments for plot-type, color, title, size of text, color-scheme, and toggle log transformation for the x and y axis.

3.  `categorical_eda:` This function takes in a data frame object and one categorical feature, to produce a histogram plot that visualizes the distribution of the feature. Users can also choose to plot the density graph of the feature by specifying in plot\_type. The function also offers customization on color, plot title, font size, color-scheme, plot size, opacity level, and facet factor.

This document will walk through an example problem to increase the understanding of the role of these functions to generate 3 types of exploratory data analysis plots.

## Install and Load the Package

You can download, build and install this package from GitHub with:

`install.packages("devtools")`

`devtools::install_github("UBC-MDS/RSimplerEDA")`

You can load the package with:

```{r load_package}

library(RSimplerEDA)
library(magrittr)

```

## Data: palmerpenguins

This example uses an open source dataset known as `penguins` from [palmerpenguins](https://allisonhorst.github.io/palmerpenguins/) package to demonstrate the application of our RSimplerEDA package. There is eight features in the penguins dataset: `species`, `island`, `bill_length_mm`, `bill_depth_mm`, `flipper_length_mm`, `body_mass_g`, `sex` and `year`. Penguins automatically loads as a `tibble` after being imported, but if your data is not a dataframe (e.g. matrix, csv, etc.), you will need to convert it to a data.frame or tibble before using the RSimplerEDA package.

```{r load data}

library(palmerpenguins)
dim(penguins)
is.data.frame(penguins)
penguins

```

## Using RSimplerEDA for Exploratory Data Analysis Plots

There are 344 observations in the `Penguins` dataset with the features.

Questions that might arise when you want to explore the data:

-   Is there correlation among numerical variables in the dataset? (e.g between the `species` and `body_mass_g` or among all numerical variables)

-   Is there linear relationship between two numerical variables? (e.g. `bill_length_mm` and `body_mass_g`)

-   What is the distribution of `body_mass_g` in different `island`?

This package looks at answering the questions by providing common plots to explore and visualize the different features and their relationships.

### Handling Missing Values

If the data you use has missing values, you should handle it in the manner that fit your objective (e.g imputation or drop all of them). In the following example, we will drop all NAs in the penguins data so that our package will work properly.

```{r Handling Missing Values}

penguins_drop_na <- penguins %>% tidyr::drop_na()
```

### 1. Find the Correlation among Numeric Variables (`corr_map`)

The function `coor_map` takes in dataframe and a character vector with numerical variables, and produce a correlation heatmap with the correlation coefficients of -1 to 1 among the given numerical variables. Users are allowed to set multiple arguments regarding the setting of the correlation plot, including method to calculate the correlation, color schemes and plot title.

Identifying correlation is important in the EDA process. It not only provides an overview to the relationship among numerical variables, it can help identify features that are correlated with the target if we want to conduct regression analysis.\

In the example questions regarding correlation among numerical variables in the penguin dataset, there seems to be high positive correlation between flipper length and body mass

```{r corr_map, fig.height=4, fig.width=7}


corr_map(penguins_drop_na, c("bill_length_mm", "bill_depth_mm", "flipper_length_mm", "body_mass_g"))
```

### 2. Finding the relationship of two Numeric Variables (`numerical_eda`)

This function takes in a data frame object, two numeric columns, and produces either a "scatter" or "line" plot to visualize the relationship between the two numerical features. Users can optionally change default arguments for plot-type, title, size of text, color-scheme, and toggle log transformation for the x and y axis. In the penguins example, we are interested to know whether higher `body_mass_g` is associated to longer `bill_length_mm` for different species. Hence we will use `numerical_eda` to plot this association.

```{r numerical_eda, message=FALSE, warning=FALSE, fig.height=4, fig.width=6}

numerical_eda(penguins, body_mass_g, bill_length_mm, color = species, title = "Body mass (grams) vs Bill Length (mm)")
```

Optionally we can also look at the relationship of `body_mass_g` and `bill_length_mm` based on the `sex` of penguins.

```{r numerical_eda log tranform, message=FALSE, warning=FALSE, fig.height=4, fig.width=6}

numerical_eda(penguins, body_mass_g, bill_length_mm, color = sex, title = "Body mass (grams) vs Bill Length (mm)")
```

### 3. Visualize the distribution of `body_mass_g` in different `island` (`categorical_eda`)

`categorical_eda` takes in data frame as input data, a categorical feature as color, and a numerical or categorical feature as value on x-axis, to produce a histogram plot that visualize the distribution of the numerical feature on each categorical variable. With `penguins` data frame as an example, `categorical_eda` plots three faceted histogram that shows the distribution of `body_mass_g` of penguins from different island.

```{r categorical_eda histogram, warning = FALSE, fig.height=4, fig.width=6}

categorical_eda(penguins, 
                xval = body_mass_g, 
                plot_type = "histogram", 
                color=island, 
                facet_factor = "island", 
                facet_col = 1, 
                title="Distribution of Body Mass of Penguins from Each Island", 
                font_size = 8)
```

In addition to "histogram", `categorical_eda` also has the option to plot density graph by changing plot\_type to density and adding an opacity value.

```{r categorical_eda density, warning = FALSE, fig.height=4, fig.width=6}

categorical_eda(penguins, 
                xval = body_mass_g, 
                plot_type = "density", 
                color=island, 
                opacity = 0.6, 
                title="Density Graph of Body Mass of Penguins from Each Island", 
                font_size=8)
```
